---
title: "LineageInference"
output: html_notebook
---

Seurat running and integration
```{r}
library(Seurat)
library(Matrix)
load("/data/proj/GCB_DvB/humanPublication/pubsplicedfull.Rdata")
load("/data/proj/GCB_DvB/humanPublication/pubunsplicedfull.Rdata")
load("/data/proj/GCB_DvB/humanPublication/networkAnnoFull.Rdata")
metadata <- networkAnnotableFile
```
Load original clustering
```{r}
load("/data/proj/GCB_DvB/humanPublication/clusteringmanuscipt.Rdata")
annotationpaper <- networkAnnotableFile
annotationpaper$OriginalClustering <- networkAnnotableFile$seurat_clusters
library(viridis)
library(hues)
sidecols <- iwanthue(length(unique(annotationpaper$OriginalClustering)), 0, 360, 36, 180, 13, 73)
sidecols <- sidecols[as.factor(annotationpaper$OriginalClustering)]
annotationpaper$col <- sidecols
#color matching
colorscheme <- annotationpaper$col
for(i in 1:length(levels(annotationpaper$OriginalClustering))){
  #print(i)
  clsel <- levels(annotationpaper$seurat_clusters)[i]
  colordist <- as.data.frame(table(annotationpaper$col[annotationpaper$seurat_clusters %in% clsel]))
  coloruse <- colordist$Freq
  names(coloruse) <- colordist$Var1
  coloruse <- names(sort(coloruse,decreasing = TRUE))[1]
  colorscheme[which(annotationpaper$seurat_clusters %in% clsel)] <- coloruse
}
annotationpaper$col.translated <- colorscheme
library(plyr)
Level1  <- revalue(as.factor(annotationpaper$OriginalClustering), c(
  "26"="OPC",
  "15"="Glioblast",
  "7"="Glioblast",
  "8"="Glioblast",
  "25"="Glioblast",
  "27"="VLMC",
  "23"="NeuralProgenitor",
  "14"="NeuralProgenitor",
  "6"="NeuralProgenitor",
  "3"="NeuralProgenitor",
  "13"="NeuralProgenitor",
  "17"="NeuralProgenitor",
  "11"="RadialGlia",
  "1"="RadialGlia",
  "4"="RadialGlia",
  "21"="RadialGlia",
  "20"="RadialGlia",
  "18"="RadialGlia",
  "24"="RadialGlia",
  "12"="EarlyInhibitory",
  "5"="EarlyInhibitory",
  "22"="EarlyInhibitory",
  "19"="EarlyInhibitory",
  "2"="EarlyInhibitory",
  "10"="EarlyInhibitory",
  "28"="EarlyInhibitory",
  "9"="EarlyExcitatory",
  "16"="EarlyExcitatory",
  "0"="EarlyExcitatory",
  "29"="Unknown",
  "30"="Unknown",
  "31"="Unknown"
  ))
Level1 <- factor(Level1, levels= c(
  "OPC",
  "Glioblast",
  "RadialGlia",
  "NeuralProgenitor",
  "EarlyInhibitory",
  "EarlyExcitatory",
  "VLMC",
  "Unknown"))
levels(Level1)
c("OPC","Glioblast","RadialGlia","NeuralProgenitor","EarlyInhibitory","EarlyExcitatory","VLMC","29","30","31")     
Level1colpal <- c("#E8C688","#D3AC9C","#B9CB6C","#738BC6","#D0D1BD","#DA8B6C","#E0C196","#C9B9A4","#C9B9A4","#C9B9A4") 
Level1col <- Level1colpal[as.factor(Level1)]

annotationpaper$Age <- annotationpaper$Sample
library(plyr)
annotationpaper$Age  <- revalue(as.factor(annotationpaper$Age), c(
  "hF_WK10_2017"="PCW 10",
  "hF_WK10psaNCAM_2017"="PCW 10",
  "hF_WK8_2017"="PCW 8",
  "hF_WK8.5cer.ctx_2019"="PCW 8.5",
  "hF_WK8.5fb_2019"="PCW 8.5",
  "hF_WK9_2019"="PCW 9"))
annotationpaper$Age <- factor(annotationpaper$Age, levels= c(
  "PCW 8",
  "PCW 8.5",
  "PCW 9",
  "PCW 10"))
#merge psaNCAM with non-psaNCAM
annotationpaper$Samples  <- revalue(as.factor(annotationpaper$Sample), c(
   "hF_WK10_2017"="hF_WK10V2",
  "hF_WK10psaNCAM_2017"="hF_WK10V2",
  "hF_WK8_2017"="hF_WK8V2",
  "hF_WK8.5cer.ctx_2019"="hF_WK8.5cer.ctxV3",
  "hF_WK8.5fb_2019"="hF_WK8.5fbV3",
  "hF_WK9_2019"="hF_WK9V3"))

annotationpaper$Level1 <- Level1
annotationpaper$Level1col <- Level1col
colorscheme <- annotationpaper$Level1col
for(i in 1:length(levels(annotationpaper$Level1))){
  #print(i)
  clsel <- levels(annotationpaper$seurat_clusters)[i]
  colordist <- as.data.frame(table(annotationpaper$Level1col[annotationpaper$seurat_clusters %in% clsel]))
  coloruse <- colordist$Freq
  names(coloruse) <- colordist$Var1
  coloruse <- names(sort(coloruse,decreasing = TRUE))[1]
  colorscheme[which(annotationpaper$seurat_clusters %in% clsel)] <- coloruse
}
annotationpaper$Level1col.translated <- colorscheme
colorscheme <- annotationpaper$Level1
for(i in 1:length(levels(annotationpaper$seurat_clusters))){
  #print(i)
  clsel <- levels(annotationpaper$seurat_clusters)[i]
  colordist <- as.data.frame(table(annotationpaper$Level1[annotationpaper$seurat_clusters %in% clsel]))
  coloruse <- colordist$Freq
  names(coloruse) <- colordist$Var1
  coloruse <- names(sort(coloruse,decreasing = TRUE))[1]
  colorscheme[which(annotationpaper$seurat_clusters %in% clsel)] <- coloruse
}
annotationpaper$Level1.translated <- colorscheme

```
Integrate the version 2 and version 3 data using Seurat CCA
```{r}
library(Matrix)
# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# Create our Seurat object and complete the initalization steps
hF <- CreateSeuratObject(counts = emats, project = "hFetal", meta.data = metadata)
#split the dataset into a list of two seurat objects
hF.list <- SplitObject(hF, split.by = "version")
hF.list <- lapply(X = hF.list, FUN = function(x) {
    x <- subset(x, subset = nFeature_RNA > 10)
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 10000)
})
#select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = hF.list,nfeatures = 10000)
hF.anchors <- FindIntegrationAnchors(object.list = hF.list, anchor.features = features,reduction = "cca",dims = 1:30)
#this command creates an 'integrated' data assay
hF <- IntegrateData(anchorset = hF.anchors)
DefaultAssay(hF) <- "integrated"
hF@meta.data$seurat_clusters2 <- hF@meta.data$seurat_clusters
#hF <- NormalizeData(hF)
hF <- FindVariableFeatures(hF, selection.method = "vst",nfeatures = 10000)
hF <- ScaleData(hF, features = VariableFeatures(hF))
hF <- RunPCA(hF, features = VariableFeatures(hF), ndims.print = 1:5, nfeatures.print = 3)
# view cell cycle scores and phase assignments
# Visualize the distribution of cell cycle markers across
hF <- RunUMAP(hF, dims=1:50,n.neighbors = 200,n.components = 2)
hF <- CellCycleScoring(hF, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
# hF@meta.data$seurat_clusters2 <- metadata$seurat_clusters
hF <- FindNeighbors(hF, dims = 1:50,k.param = 10)
hF <- FindClusters(hF, resolution = 1)
RidgePlot(hF, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), ncol = 2)
DimPlot(hF,label = TRUE,group.by = "Phase",split.by = "Sample",ncol=2)
DimPlot(hF,label = TRUE,group.by = "Sample",split.by = "version",ncol=2)
DimPlot(hF,label = TRUE,group.by = "version",shuffle = T)
DimPlot(hF,label = TRUE,group.by = "Sample",shuffle = T)
DimPlot(hF,label = TRUE,group.by = "seurat_clusters2")
DimPlot(hF,label = TRUE,group.by = "seurat_clusters")
```
Write csv to load in python and run scvelo for transition matrix
```{r}
write.table(hF@assays$integrated@data,file = "/data/proj/GCB_DvB/humanPublication/pubsplicedcorrectedDM.csv",sep=",")
write.table(hF@meta.data,file = "/data/proj/GCB_DvB/humanPublication/pubannoDM.csv",sep=",")
#These are not needed
write.table(hF@reductions$umap@cell.embeddings,file = "/data/proj/GCB_DvB/humanPublication/umapscVeloDM.csv",sep=",")
ematu <- ematu[,colnames(emats)]
# row.names(emats) <- row.names(hF)
# row.names(ematu) <- row.names(hF)
write.table(emats[row.names(hF@assays$integrated@data),colnames(hF)],file = "/data/proj/GCB_DvB/humanPublication/pubsplicedFtruncDM.csv",sep=",")
write.table(ematu[row.names(hF@assays$integrated@data),colnames(hF)],file = "/data/proj/GCB_DvB/humanPublication/pubunsplicedFtruncDM.csv",sep=",")
#load("/data/proj/GCB_DvB/humanPublication/Final/NBpart1.RData")
```
After running scVELO in python we load the results back in.
```{r}
library(Matrix)
library(Seurat)
#load("/data/proj/GCB_DvB/humanPublication/Step1vanBruggen5nnCCAFINAL2.Rdata")
load("/data/proj/GCB_DvB/humanPublication/pubsplicedfull.Rdata")
load("/data/proj/GCB_DvB/humanPublication/pubunsplicedfull.Rdata")
emats <- emats#[,dup.cells]
ematu <- ematu#[,dup.cells]
scVtm <- readMM('/data/proj/GCB_DvB/humanPublication/scVELOtransmatFSCCA30NNveloDM.mtx')
lt <- read.csv("/data/proj/GCB_DvB/humanPublication/scVELOLT_FSCCA30NNveloDM.csv")
#Optional read in transcription factor annotations from SCENIC
TFsSCENIC <- read.table(file="/data/proj/GCB_DvB/human/hFetal/motifs-v9-nr.hgnc-m0.001-o0.0.txt",sep="\t",stringsAsFactors = F)
TFs <- unique(TFsSCENIC$V6)
```
Now we create the diffusion map integrating both velocity and transcriptional distances into one manifold
```{r}
set.seed(1234)
#First we create the diffusion map based on transcription
tsnerlba <- t(hF@reductions$pca@cell.embeddings[,1:50])
#get destiny, removed from bioconductor (not very useful >:())
# install.packages("remotes")
# remotes::install_github("https://github.com/theislab/destiny",build_vignettes = FALSE, dependencies = TRUE)
#check for duplicated cells 
# dup.cells <- !duplicated(t(tsnerlba))
# table(dup.cells)
# dup.cells <- as.logical(dup.cells)
# hF <- hF[,dup.cells]
# tsnerlba <- tsnerlba[,dup.cells]

library(destiny)
tsmanifolddist <- DiffusionMap(data =t(tsnerlba) ,n_eigs =20,verbose = TRUE)#,sigma=0.001) 
colnames(tsmanifolddist@transitions) <- colnames(hF)
row.names(tsmanifolddist@transitions) <- colnames(hF)

#Now we create the joint representation
XY <- t(scVtm) %*% tsmanifolddist@transitions

#Now we integrate the spaces
library(irlba)
dcompl <- irlba(XY,50,maxit = 100000)
tsnerlba <- t(dcompl$u)
tsnerlba <- rbind(tsnerlba,t(dcompl$v))
library(destiny)
tsmanifold <- DiffusionMap(data =t(tsnerlba) ,n_eigs =20,verbose = TRUE)#,sigma=0.001) 
colnames(tsmanifold@transitions) <- colnames(hF)
row.names(tsmanifold@transitions) <- colnames(hF)

#Now we generate the backward kernel 80% velocity 20% transcriptional distance
scVtmb <-  scVtm
scVtmb <- t(t(scVtmb)/Matrix::colSums(scVtmb,na.rm = TRUE))
scVtmb[is.na(scVtmb)] <- 0
kernel <- t((scVtm *0.8)) + (tsmanifold@transitions * 0.2)
kernel <- t(t(kernel)/Matrix::colSums(kernel,na.rm = TRUE))
kernel[is.na(kernel)] <- 0
```
Now we calculate the new clusters based on the new manifold
```{r}
# library(Seurat)
# #library(amap)
#77
newclustobj <- Seurat::CreateSeuratObject(counts = as.matrix(hF@assays$RNA@data))
NetworkDist <- as.dist(1-tsmanifold@transitions)
newclustobj[["HSphere_snn"]] <- Seurat::FindNeighbors(NetworkDist,k.param=10)$snn
newclustobj <- Seurat::FindClusters(newclustobj, graph.name = "HSphere_snn",resolution = 1.6) #algorithm = 4,
table(newclustobj@meta.data$seurat_clusters)
hF@meta.data$seurat_clusters3 <- newclustobj@meta.data$seurat_clusters
DimPlot(hF,label = TRUE,group.by = "seurat_clusters3")+ NoLegend()
Idents(hF) <- 'seurat_clusters3'
  DimPlot(hF, label=F, cells.highlight=WhichCells(hF, idents = "69"), cols.highlight = c("darkblue", "darkred"), cols= "grey")
  FeaturePlot(hF,features = c("APOD"),slot="data")
```

And we calculate the pseudotimes associated to each cluster
```{r}
pt <- destiny::DPT(tsmanifold)#,tips = c(1,2))
cd_diffusionplot <- hF@meta.data
lineagecluster <- as.factor(hF$seurat_clusters3)
names(lineagecluster) <- row.names(cd_diffusionplot)
x<- pt$DPT1
names(x) <- row.names(cd_diffusionplot)
branchtips = as.character(NULL)#lowestcell
branchpseudotimes <- matrix(nrow=(length(levels(as.factor(lineagecluster)))),ncol=nrow(cd_diffusionplot))
lineagecluster <- lineagecluster[row.names(cd_diffusionplot)]
for(i in 1:length(levels(droplevels(as.factor(lineagecluster))))){
  pseudotime <- x[names(lineagecluster[lineagecluster %in% levels(lineagecluster)[i]])]
  topcell <- names(pseudotime[pseudotime == max(pseudotime)])
  branchtips <-  c(branchtips,topcell)
x <-pt[,which(row.names(cd_diffusionplot)==topcell,arr.ind = TRUE)]
names(x) <- row.names(cd_diffusionplot)
  branchpseudotimes[i,] <- x
  #print(i)
}
colnames(branchpseudotimes) <- row.names(cd_diffusionplot)
row.names(branchpseudotimes) <- levels(as.factor(lineagecluster))
branchpseudotimes <- branchpseudotimes[as.character(sort(as.integer(row.names(branchpseudotimes)),decreasing = F)),]

x <- hF@reductions$umap@cell.embeddings[,1]
  y <- hF@reductions$umap@cell.embeddings[,2]
  
load("/data/proj/GCB_DvB/humanPublication/umapharmony.Rdata")
  x <- umap_out_harmony2ddist$layout[,1]
  y <- umap_out_harmony2ddist$layout[,2]
```
Now we add tracks of lineage diffusion giving the start states of the diffusion process #this is the one to use
```{r}
#library(e1071)
library(Matrix)
library(destiny)
startpt <- lt$latent_time
names(startpt) <- colnames(branchpseudotimes)
traceclusters <- matrix(ncol=nrow(branchpseudotimes),nrow=ncol(branchpseudotimes))
for(i in 1:nrow(branchpseudotimes)){
  #print(i)
log<- numeric()
starttime = "-1"
if(!starttime == row.names(branchpseudotimes)[i]){
source <- startpt
sink <- branchpseudotimes[i,]
sink <- sort(sink)
trackcellstart <- names(head(sink,10))
lineagecells <- trackcellstart
trackcell <- trackcellstart[1]
threshold <- source[trackcellstart[1]]
stall = 1
while(source[trackcell] > 0 & stall > 0 ) {
  logold <- log
  log<- c(log,source[trackcellstart])
  stall = length(log) - length(logold)
pool <- names(head(sink[sink > sink[trackcell]],1))
if(length(pool) < 1) {trackcell <- names(which(source == min(source)))}
else{
trackcellstart <- names(which(source[pool] <= min(log)))
lineagecells <- c(lineagecells,trackcellstart)
trackcell <- names(which(source[pool] == min(source[pool])))}
}
LineageCells <- lineagecells
Pathtrace <- names(lineagecluster) %in% c(unique(LineageCells))#,unique(LineageCells2))
}
else{Pathtrace <- logical(ncol(branchpseudotimes)) }
traceclusters[,i] <- Pathtrace
}
```
Here we start the diffusion process, first add the function
```{r}
WAdiffuse <- function(A,dist,ncells,pool=FALSE,cores,distselectneighbor,nameselectneighbor){
  library(Matrix)
  library(foreach)
  library(doSNOW)
  library(matrixStats)
  range01 <- function(x){(x-min(x))/(max(x)-min(x))}
 cl <- makeCluster(cores)  
 registerDoSNOW(cl)  
iterations <- ncol(A)
pb <- txtProgressBar(max = iterations, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)
#distselectneighbor <- Matrix(apply(dist,2,function(x) sort(x,decreasing=FALSE)[1:ncells]),sparse = TRUE)
#nameselectneighbor <- apply(dist,2,function(x) sort(x,decreasing=FALSE,index.return=TRUE)$ix[1:ncells])
rm(dist)
gc()
if(class(A) != "dgCMatrix"){
A <- Matrix(as.matrix(A),sparse = T)
}
gc()
if(pool==FALSE){
D <-  foreach(i = 1:ncol(A),.combine=cbind,.packages=c("Matrix"),.options.snow = opts) %dopar% {
distsub <- 1-distselectneighbor[,i] #uses prob
C <- A[,nameselectneighbor[,i]]
C <- apply(C,1,function(x) weighted.mean(x,distsub))
return(C)
}
close(pb)
 stopCluster(cl)
}
if(pool==TRUE){
D <-  foreach(i = 1:ncol(A),.combine=cbind,.packages=c("Matrix"),.options.snow = opts) %dopar% {
distsub <- rowSums(A[,nameselectneighbor[,i]])
return(distsub)
}
close(pb)
 stopCluster(cl)
}
colnames(D) <- colnames(A)
row.names(D) <- row.names(A)
D <- Matrix(D,sparse=TRUE)
  gc()
 return(D)
}
```
Now we run the diffusion for 30 iterations with 100 neighbors, and calculate lineage membership and diffusion score

```{r}
library(Matrix)
nameselectneighbor <- apply(kernel,2,function(x) sort(x,decreasing=TRUE,index.return=TRUE)$ix[1:100])
distselectneighbor <- Matrix(apply(kernel,2,function(x) sort(x,decreasing=TRUE)[1:100]),sparse = TRUE)

linwalk <- list()
linwalk[[1]] <- WAdiffuse(t(1*traceclusters),kernel,pool = FALSE,ncells=100,cores=8,distselectneighbor,nameselectneighbor)
for(i in 2:30){
  print(i)
linwalk[[i]] <- WAdiffuse(linwalk[[i-1]],kernel,pool = FALSE,ncells=100,cores=8,distselectneighbor,nameselectneighbor)
}

#save(linwalk,file="/data/proj/GCB_DvB/humanPublication/LinwalkFTRUECCAdiff30NNusingkernel.Rdata")
#load("/data/proj/GCB_DvB/humanPublication/LinwalkFTRUECCAdiff30NNusingkernel.Rdata")

#get lineages
for(j in 1:nrow(linwalk[[1]])){
  fpass <- linwalk[[1]][j,]
for(i in 2:30){
fpass <- cbind(fpass,linwalk[[i]][j,])
}
values <- apply(fpass,1,max)
if(j == 1){lineage = values} else{lineage <- rbind(lineage,values)}
}
```
Get lineage KL function
```{r}
KLlin <- function(A,B,cores,skipgenes,limitcells,ncells,correct=FALSE,batch){
  print(correct)
  library(Matrix)
  library(foreach)
  library(doSNOW)
  range01 <- function(x){(x-min(x))/(max(x)-min(x))}
 cl <- makeCluster(cores)  
 registerDoSNOW(cl)  
iterations <- ncol(A)
pb <- txtProgressBar(max = iterations, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)
if(class(A) != "dgCMatrix"){
A <- Matrix(as.matrix(A),sparse = T)
 }
 if(class(B) != "dgCMatrix"){
B <- Matrix(as.matrix(A),sparse = T)
 }
if(min(A) < 0){
A <- A-min(A)
}
if(min(B) < 0){
B <- B-min(B)
}
rm(B)
gc()
D <-  foreach(i = 1:ncol(A),.combine=cbind,.packages="Matrix",.options.snow = opts) %dopar% {
D <- rep(NA, ncol(A))
if(skipgenes==TRUE){
avoidgene <- which(! (A[,i]<=0));C <- A[avoidgene,]}else{C <- A}
Ckl   <- t(t(C)/colSums(C))#[avoidgene,]
Ckl[is.na(Ckl)] <- 0
Ckl <- Matrix(Ckl+0.0000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000001,sparse=TRUE)
Ckl_log <- log(Ckl[,i]/Ckl[,-i])
C1kl <- Ckl[,i]*Ckl_log
rm(Ckl_log)
Ckl <- colSums(C1kl)

nnsort <- sort(Ckl,decreasing = FALSE)

if(limitcells==TRUE){
G <- t(C1kl[,names(nnsort)[1:ncells]]) / Ckl[names(nnsort)[1:ncells]];G <- (t(G) %*% Ckl[names(nnsort)[1:ncells]])[,1]}else{G <- t(C1kl) / Ckl;G <- (t(G) %*% Ckl)[,1]}

GD <- rep(NA, nrow(A))
usedG <- which(row.names(A) %in% names(G))
notusedG <- which(!row.names(A) %in% names(G))
GD[usedG] <- G
GD[notusedG] <- 0
names(GD) <- row.names(A)

C <- GD

gc()
return(C)
}
close(pb)
 stopCluster(cl)
 
colnames(D) <- colnames(A)
rownames(D) <- row.names(A)
Dnorm <- scale(abs(D), center=FALSE, scale=colSums(abs(D)))
Dnorm.sign <- scale(D, center=FALSE, scale=colSums(abs(D)))
D <- Matrix(D,sparse=TRUE)
Dnorm <- Matrix(Dnorm,sparse=TRUE)
Dnorm.sign <- Matrix(Dnorm.sign,sparse=TRUE)
D <- list(D,Dnorm,Dnorm.sign)
names(D) <- c("D","Dnorm","Dnorm.sign")
  #D <- D/sum(D)
  gc()
 return(D)
}
```
Run lineage membership estimation for each cell in the matrix
```{r}
lineageKL <- lineage
lineageKLdone <- KLlin(lineage,lineage,15,skipgenes = FALSE,limitcells = FALSE,ncells = 100,correct = FALSE,batch = colnames(hF))
#save(lineageKL,file="/data/proj/GCB_DvB/humanPublication/Final/lineagKLfinalrealCCA30NNvelokernel.Rdata")
#load("/data/proj/GCB_DvB/humanPublication/Final/lineagKLfinalrealCCA30NNvelokernel.Rdata")
lineage.mem <- lineageKLdone$Dnorm
lineage.mem.max <- (apply(lineage.mem,2,max))

```
Visualize lineage information
```{r}
library(plotly)
library(dplyr)
#Differentiation score
plot_ly( x = ~x, y = ~y,marker = list(size = 5)) %>% add_markers(color=~lineage.mem.max,alpha=1)

#Plot OL lineage cluster 69 , row 70.
Expression_capped <- lineage.mem[70,]^0.1
Expression_capped <- scale(Expression_capped)
Expression_capped[Expression_capped <= 1.96]  <- 0
    names(Expression_capped) <- seq_len(length(Expression_capped))
    Expression_capped <- sort(Expression_capped,decreasing = FALSE)
    ord <- as.integer(names(Expression_capped))
    plot_ly( x = ~x[ord], y = ~y[ord],marker = list(size = 5)) %>% add_markers(color=~Expression_capped,alpha=1)%>% layout(
      xaxis = list( showgrid = FALSE,showaxis=FALSE,zeroline=FALSE),
      yaxis = list(showgrid = FALSE,showaxis=FALSE,zeroline=FALSE))
```

```{r fig.height=10, fig.width=10}
library(Matrix)
emat.norm <- hF@assays$RNA@data
#REMEMBER its PLUS1 for the clusters
lnum <- as.numeric(70)
Expression_capped <- lineage.mem[lnum,]^0.1
Expression_capped <- scale(Expression_capped)
#1.645 10% 1.96 5% 2.58 1%  
Expression_capped[Expression_capped <= 1.96]  <- 0
ptlin <- which(Expression_capped > 0)
length(ptlin)

Lineagegenes <- row.names(hF@assays$integrated@data)
#Only TFs
#Lineagegenes <- Lineagegenes[Lineagegenes %in% TFs]
suppressWarnings(corlin <- apply(as.matrix(emat.norm[Lineagegenes,ptlin]),1,function(x) cor.test(x,lineage.mem[lnum,ptlin]^0.1,method="spearman")))
linKLdrivers <- as.data.frame(t(as.data.frame(lapply(corlin,unlist))))
linKLdrivers$p.value <- as.numeric(as.character(linKLdrivers$p.value))
linKLdrivers$estimate.rho <- as.numeric(as.character(linKLdrivers$estimate.rho))
#linKLdrivers$p.adjust <- p.adjust(linKLdrivers$p.value,method = "BH")
ldsig <- linKLdrivers$p.value;ldmaxcor <-linKLdrivers$estimate.rho
names(ldmaxcor) <- Lineagegenes
names(ldsig) <- Lineagegenes
lineagevals <- as.data.frame(cbind(ldsig,ldmaxcor))
lineagevals <- subset(lineagevals,lineagevals$ldsig < 0.05)
lvgenes <- lineagevals
Lineagegenes <- row.names(lineagevals[order(abs(lineagevals$ldmaxcor),decreasing = TRUE),])#[1:500]


linclusters <- as.data.frame(table(hF$seurat_clusters3[ptlin])/table(hF$seurat_clusters3))
linclusters <- as.character(linclusters[linclusters$Freq >0.1,]$Var1)
corclusters <- hF$seurat_clusters3#[ptlin]
 for(i in 1:length(linclusters)){
  if(i==1){cat("Correlating genes...binning by clusters...")}
  cat(paste("Bin Cluster ",linclusters[i],"..."))
  if(i!=1){linKLdriverssig <-cbind(linKLdriverssig,linKLdrivers$p.value);linKLdriverscor <-cbind(linKLdriverscor,linKLdrivers$estimate.cor)}
binlin  <- emat.norm[Lineagegenes,]
binlin  <- binlin[,corclusters %in% linclusters[i]]
cat(paste(length(which(corclusters %in% linclusters[i])),"cells..."))
  linbin <- lineage.mem[lnum,]^0.1
  linbin <- linbin[corclusters %in% linclusters[i]]
suppressWarnings(corlin <- apply(as.matrix(binlin),1,function(x) cor.test(x,linbin,method="pearson")))
linKLdrivers <- as.data.frame(t(as.data.frame(lapply(corlin,unlist))))
linKLdrivers$p.value <- as.numeric(as.character(linKLdrivers$p.value))
linKLdrivers$p.value <- p.adjust(linKLdrivers$p.value,method = "BH")
linKLdrivers$estimate.cor <- as.numeric(as.character(linKLdrivers$estimate.cor))
if(i!=1){linKLdriverssig <-cbind(linKLdriverssig,linKLdrivers$p.value)}
if(i==1){linKLdriverssig <- linKLdrivers$p.value;linKLdriverscor <-linKLdrivers$estimate.cor}
 }
row.names(linKLdriverscor) <- Lineagegenes
row.names(linKLdriverssig) <- Lineagegenes
ldmaxcor <- apply(linKLdriverscor,1,max,na.rm = TRUE)
ldmaxvar <- apply(linKLdriverscor,1,var,na.rm = TRUE)
names(ldmaxcor) <- Lineagegenes
names(ldmaxvar) <- Lineagegenes
ldsig <- apply(linKLdriverssig,1,min,na.rm = TRUE)
names(ldsig) <- Lineagegenes
lineagevals <- as.data.frame(cbind(ldsig,ldmaxcor,ldmaxvar))
lineagevals <- subset(lineagevals,lineagevals$ldsig < 0.05)
lvgenes <- lineagevals
EC <- lineage[lnum,ptlin]
names(EC) <- seq_len(length(EC))
EC <- sort(EC,decreasing = FALSE)
ord <- as.integer(names(EC))

if(nrow(lineagevals) <=100){ptgenes <- row.names(lineagevals)}else{ptgenes <- row.names(lineagevals[order(lineagevals$ldmaxcor,decreasing = TRUE),])[1:100]}

ctlin <- as.matrix(emat.norm[ptgenes,ptlin])
dat <- data.frame(t(ctlin[,ord]))
## response
vars <- colnames(dat)
## covariate
id <- 1:nrow(dat)
library(mgcv)
dat <- data.frame(t(ctlin[,ord]))
colnames(dat) <- gsub("-", "", colnames(dat))
ptgenesnum <- colnames(dat)
dat <- cbind(id,dat)

new.datp <- matrix(nrow=length(ptgenesnum),ncol=length(id))
new.datpval <- numeric(length=length(ptgenesnum))
for(i in 1:length(ptgenesnum)){
  form <- as.formula(paste0(ptgenesnum[i],"~",paste0("s(","id",',bs="cs",k=10',")",collapse="+")))
  m <- gam(form, data=dat)
  form <- as.formula(paste0(ptgenesnum[i],"~",paste0("1",collapse="+")))
  m0 <- gam(form, data=dat)
  new.datp[i,] <- m$fitted.values
}
row.names(new.datp) <- ptgenes
colnames(new.datp) <- colnames(ctlin[,ord])
new.datp[new.datp<0] <- 0
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
new.datp <- t(apply(new.datp,1,range01))
new.dat <- new.datp#[fdr < 0.05,]

GC <- unlist(apply(new.dat,1,function(x) which(x == max(x))[1]))
names(GC) <- seq_len(length(GC))
GC <- sort(GC,decreasing = TRUE)
ordG <- as.integer(names(GC))

ctlin <- as.matrix(ematu[row.names(new.dat),ptlin])
dat <- data.frame(t(ctlin[,ord]))
vars <- colnames(dat)
id <- 1:nrow(dat)
library(mgcv)
dat <- data.frame(t(ctlin[,ord]))
colnames(dat) <- gsub("-", "", colnames(dat))
ptgenesnum <- colnames(dat)
dat <- cbind(id,dat)
new.datu <- matrix(nrow=length(ptgenesnum),ncol=length(id))
for(i in 1:length(ptgenesnum)){
  #print(i)
  form <- as.formula(paste0(ptgenesnum[i],"~",paste0("s(","id",',bs="cs",k=10',")",collapse="+")))
  new.datu[i,] <- gam(form, data=dat)$fitted.values
}
row.names(new.datu) <- row.names(new.dat)
colnames(new.datu) <- colnames(ctlin[,ord])
new.datu[new.datu<0] <- 0
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
new.datu <- t(apply(new.datu,1,range01))
new.dat[new.dat<0] <- 0
new.datu <- new.dat[ordG,]-new.datu[ordG,]
filter <- apply(new.datu,1,anyNA)
new.datu <- new.datu[!filter,]
ordF <- ordG[!filter]
GC <- unlist(apply(new.datu,1,function(x) which(x == max(x))[1]))
names(GC) <- seq_len(length(GC))
GC <- sort(GC,decreasing = TRUE)
ordF <- as.integer(names(GC))


library(heatmap3)
library(pheatmap)
library(hues)
library(viridis)
sidecols <- iwanthue(length(unique(hF@meta.data$seurat_clusters3)),0, 360, 36, 180, 13, 73)
sidecols <- sidecols[as.factor(hF@meta.data$seurat_clusters3)][ptlin]
names(sidecols) <- seq_len(length(sidecols))
sidecols <- sidecols[ord]
heatmap3(new.dat[ordG,], Rowv = NA, Colv = NA ,scale = "none",symm = F, method = "ward.D2",col=viridis(1024),balanceColor =F,cexRow = 1.2,cexCol = 0.5,ColSideColors=sidecols,labCol = NA,)
heatmap3(new.datu[ordF,], Rowv = NA, Colv = NA ,scale = "none",symm = F, method = "ward.D2",balanceColor =F,cexRow = 1.4,cexCol = 0.5,ColSideColors=sidecols,labCol = NA)

save.image("/data/proj/GCB_DvB/humanPublication/Final/NBpart2.RData")
```